### 2023-09-12 language-v4/proofs/optimizations/CaseOfKnownConstructor.tex

\CaseOfKnownConstructorTheorem

\begin{proof}
    Trivial through the progress theorem.
\end{proof}