### 2023-09-12 language-v4/proofs/optimizations/Inlining.tex

\InliningTheorem

\begin{proof}~

\begin{tabbing}
    (1) $\G,\G',\D \vdash \llet{\xD = e}{e'} : \vp$\\
    (2) $\G,\D,\xD \vdash e' : \vp$\` by inv. on (let)\\
    (3) $\G',\D \vdash e : \s$\` by inv. on (let)\\
    (4) $\hasnolinearvars{\G'}$\\
    (5) $\G,\G',\D\vdash e'[e/x] : \vp$\` by $\D$-subst. lemma (2,3)\\
    (6) $\G,\G',\D,\xD \vdash e'[e/x] : \vp$ \` by $Weaken_\Delta$\\
    (7) $\G,\G',\G',\D \vdash \llet{\xD = e}{e'[e/x]} : \vp$\` by (let) (7,3,4)\\
    (8) $\G,\G',\D \vdash \llet{\xD = e}{e'[e/x]} : \vp$\` by (4)\\
\end{tabbing}
\end{proof}