### 2023-09-13 language-v4/proofs/optimizations/CaseOfKnownConstructor.tex
Write and typeset multiple optimisation proofs

\CaseOfKnownConstructorTheorem

\begin{proof}~
\begin{tabbing}
    (1) $\G; \D, \D' \vdash \ccase{K~\ov{e}}{\z[\D][\s]~\{..., K~\ov{x} \to e_i\}} : \vp$\\
    (2) $\ccase{K~\ov{e}}{\z[\D][\s]~\{..., K~\ov{x} \to e_i\}} \longrightarrow e_i\ov{[e/x]}[K~\ov{e}/z]$\\
    (3) $\G; \D,\D' \vdash e_i\ov{[e/x]}[K~\ov{e}/z] : \vp$\` by preservation theorem (1,2)\\
\end{tabbing}
\end{proof}