### 2023-09-22 language-v4/proofs/optimizations/CaseOfKnownConstructor.tex

Case-of-known constructor is a transformation that essentially evaluates a case
expression of a known constructor at compile time, by substituting in the
matching alternative the case binder by the scrutinee and pattern variables by
constructor arguments.
%
Intuitively, either the case binder is used exactly once to consume the
resources of the scrutinee, or the pattern components whose matching
constructor argument uses linear resources are used exactly once, so one of the
substitutions is a no-op.  If the pattern variables are substituted by the
matching scrutinee expressions, the expressions are still only used once, and
if the case binder is substituted by the scrutinee, it is still used exactly
once.
%
The proof follows trivially from the preservation theorem.

\CaseOfKnownConstructorTheorem

\begin{proof}~
\begin{tabbing}
    (1) $\G; \D, \D' \vdash \ccase{K~\ov{e}}{\z[\D][\s]~\{..., K~\ov{x} \to e_i\}} : \vp$\\
    (2) $\ccase{K~\ov{e}}{\z[\D][\s]~\{..., K~\ov{x} \to e_i\}} \longrightarrow e_i\ov{[e/x]}[K~\ov{e}/z]$\\
    (3) $\G; \D,\D' \vdash e_i\ov{[e/x]}[K~\ov{e}/z] : \vp$\` by preservation theorem (1,2)\\
\end{tabbing}
\end{proof}