### 2023-10-02 prototype/core-plugin-results.tex
Make EXPERIMENT 1 about Case Var work
When run on linear-smc, with EXPERIMENT we get 0 errors, without EXPERIMENT 1 we get 4 errors that the experimente would lift. Great.

\begin{figure}[h]
\centering{
\begin{tabular}{c|c|c|c|c|c}
    \thead{Library} & \thead{Total\\Accepted} & \thead{Total\\Rejected} & \thead{Unique\\Rejected} & \thead{Case-Var\\Rejected} & \thead{Unknown\\Rejected}\\
    linear-smc & 19438 & 4 & 1 & 1 & 0\\
    linear-base & 92308 & - & - & - & -\\
\end{tabular}
}
\caption{Linear Core Plugin on Linear Libraries}
\label{fig:core-plugin-res}
\end{figure}