### 2023-10-02 prototype/core-plugin-results.tex
Corrections

\begin{figure}[h]
\centering{
\begin{tabular}{|c|c|c|c|c|c|c|c|}
    \toprule
    \thead{Library} & \thead{Total\\Accepted} & \thead{Total\\Rejected} & \thead{Unique\\Rejected} & \thead{Linear modulo\\Call-by-name} & \thead{Linear\\Rejected} & \thead{$\lnot$ Linear\\Rejected} & \thead{Unknown\\Rejected}\\
    \midrule
    linear-smc & 19438 & 4 & 1 & 1 & 0 & 0 & 0\\
    linear-base & 91516 & 499 & 72 & 10 & 8 & 2 & 52\\
    \makecell{priority-\\sesh} & 83 & 69 & 6 & - & - & - & 6\\
    \bottomrule
\end{tabular}
}
\caption{Linear Core Plugin on Linear Libraries}
\label{fig:core-plugin-res}
\end{figure}