### 2025-01-24 prototype/core-plugin-results-2.tex
started instrumented semantics

\begin{figure}[h]
\centering{
\begin{tabular}{|l|c|c|c|}
    %\toprule
    \thead{Library} & \thead{Total\\Accepted} & \thead{Total\\Rejected} & \thead{Accept Rate}\\
    %\midrule
    linear-smc & 19438 & 4 & 99.98\% \\
    \makecell{priority-sesh} & 6781 & 19 & 99.72\% \\
    linear-base & 112311 & 538 & 99.52\% \\
    %\bottomrule
\end{tabular}
}
\caption{Linear Core Plugin on Linear Libraries}
\label{fig:core-plugin-res}
\end{figure}