### 2025-11-25 prototype/core-plugin-results-2.tex
Improve plugin section with AE reviews

\begin{figure}[h]
\centering{
\begin{tabular}{|l|c|c|c|c|}
    %\toprule
    \thead{Library} & \thead{Total\\Accepted} & \thead{Total\\Rejected} & \thead{Unique\\Rejected} & \thead{Accept Rate}\\
    %\midrule
    linear-base & 48018 & 735 & 51 & 98.50\% \\
    linear-smc & 18759 & 5 & 3 & 99.97\% \\
    priority-sesh & 760 & 95 & 6 & 88.89\% \\
    text-builder-linear & 4858 & 34 & 6 & 99.30\% \\
    linear-generics & 63332 & 0 & 0 & 100.00\% \\
    %\bottomrule
\end{tabular}
}
\caption{Linear Core Plugin on Linear Libraries with GHC 9.10.3}
\label{fig:core-plugin-res}
\end{figure}